---
title: "How to extract and preprocess input data from open source"
author: "Anneke Okka Schoenfeld"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to extract and preprocess input data from open source}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Which input data is needed to run the drainage network generator?

The stormwater drainage network generator provided by the `urbandrain` package uses 
open source input data. The following data can be extracted from open source data 
bases:

- street polyline data (**streets**) for example from open street maps (OSM): 
https://www.openstreetmap.de/

- land use information (**landuse**) for example extracted from open street maps 
(OSM): https://www.openstreetmap.de/
- a digital terrain model (**dtm**) for example from the shuttle radar topography 
mission (SRTM): https://www.opendem.info/download_srtm.html

- rain timeseries (**path_raintimeseries**) for example from the Climate Data
Center provided by the DWD: https://opendata.dwd.de/climate_environment/CDC/

Additionally the follwing input data is needed. These input data require some 
basic knowledge of the model site:

- Outfall points (**outfalls**) to which the network drains to.

- A boundary polygon (**boundary_polygon**) around the outfall catchment or
catchments.

- A minimum distance value (**lim**) between street nodes (unit: meter). For 
example mean distance between inspection shafts.

- A range within the junctions dephts are allowed to vary (**min_junc_depth**, 
**mean_junc_depth**, **max_junc_depth**).

- A range within the pipe slopes are allowed to vary, based on the terrain 
(**min_slope**, **max_slope**).


## How do I extract open source data and how do I process them?

In this example data is extracted for the example case provided by this package:

```{r load_package_data}
library(urbandrain)
library(osmdata)

# access provided boundary polygon from example data:
data(ex1, package = "urbandrain")
boundary_polygon <- ex1$boundary_polygon

```

### Download street and landuse data from OSM and process them:

```{r osm_data_processing, warning=FALSE, message=FALSE}
# rectangular area for OSM extraction:
coords <- as.data.frame(sf::st_coordinates(boundary_polygon))
min_x <- min(coords$X)
max_x <- max(coords$X)
min_y <- min(coords$Y)
max_y <- max(coords$Y)

# function to extract streets, buildings and landuse from OSM:
extractFromOSM <- function(key, boundary_polygon, min_x, min_y, max_x, max_y){
  # download street data from OSM:
  data <- osmdata::opq(bbox = c(min_x, min_y, max_x, max_y))
  data <- osmdata::add_osm_feature(opq = data, key = key)
  data <- osmdata::osmdata_sf(data)
  
  # extract polylines or polygon data:
  if(key == "highway"){
    data <- data$osm_lines[c("highway", "maxspeed")]
  }
  if(key == "landuse" | key == "building"){
    data <- data$osm_polygons[key]
  }
  
  # cut to model area
  data <- sf::st_intersection(boundary_polygon, data)
  
  return(data)
  
}

# download data from OSM:
streets <- extractFromOSM("highway", boundary_polygon, min_x, min_y, max_x, max_y)
building <- extractFromOSM("building", boundary_polygon, min_x, min_y, max_x, max_y)
landuse <- extractFromOSM("landuse", boundary_polygon, min_x, min_y, max_x, max_y)

# select streets under which the drainage network shall be placed
# for example exclude motorways, paths etc.
streets_selected <- streets[streets$highway %in% c("unclassified","residential","tertiary",
                                                   "secondary","primary",
                                                   "pedestrian","trunk",
                                                   "unclassified", "living_street"),]
# ... plot data
plot(streets["highway"])


# union landuse and building data:
a <- landuse[,2]
b <- building[,2]
colnames(b)  <- c("landuse", "geometry")
op1 <- sf::st_difference(a,sf::st_union(b)) #notice the use of st_union()
op2 <- sf::st_intersection(b, a) #notice the order of b and a
op2 <- op2[,1]
colnames(op2) <- c("landuse", "geometry")
landuse_complete <- rbind(op1, op2)

# ... plot data
plot(landuse_complete)

# additionally a data.frame specifying the imperviousness of the landuse classes is needed, for example:
landuse_classes <- data.frame(class = unique(landuse_complete$landuse), 
                              imperviousness = c(0.4, 0, 0, 1, 0.5))
landuse_classes

```

### Processing of DTM data from SRTM

Download the SRTM Digital Terrain Model of Germany from https://www.opendem.info/download_srtm.html.


```{r srtm_dtm_processing, eval = TRUE}
dtm_url <- "https://www.opendem.info/downloads/srtm_germany_dtm.zip"
dtm_zip_path <- file.path(tempdir(), basename(dtm_url))
dtm_dir <- file.path(tempdir(), "srtm")

download.file(dtm_url, 
              destfile = dtm_zip_path)
archive::archive_extract(archive = dtm_zip_path, 
                         dir = dtm_dir)

# download srtm data and load data to R:
dtm_path <- normalizePath(list.files(dtm_dir, recursive = TRUE))


# load srtm tif:
dtm_germany <- raster::raster(dtm_path)

# cut raster extent to boundary_polygon:
dtm <- raster::intersect(dtm_germany, boundary_polygon)
dtm
# if needed: interpolate to get a finer raster resolution:
dtm_fine <- raster::disaggregate(dtm, fact = 10, method = "bilinear")


```

### Further parameters

To see the structure of further input parameters view ex1 example data, vignette 
[How to built a drainage network model using urbandrain](../How_to_built_a_drainage_network_model_using_urbandrain.html) 
or the documentation of the functions `create_drainage_network()` and `create_swmm_model()`.
